---
title: "Week 4 Activity"
author: 'Group 1: Yanyu LonG'
date: "9/29/2020"
output: html_document
---
  
```{r setup, include=FALSE}
setwd("E:/git/Stats506_public/activities/week4/")
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
data(Orange)
```

### Part 1 (dplyr)

In the part 1 of the activity you will write or interpret short dplyr pipes
that explore or analyze the `Orange` data from R's `datasets` package. The data
which contains 35 rows and 3 columns recording the growth
of orange trees.  The three columns in the data set are:

  - `Tree`: an ordered factor, identifying individual trees, 
  - `age`: a numeric vector giving the number of days 
    since the tree was planted,
  - `circumference`: a numeric vector recording the circumference of the 
    trunk in mm. 

You can read more at `?Orange`. 

1. Write a dplyr pipe to determine the number of
observations per tree.

```{r q1}
Orange %>% group_by(Tree) %>% count()
```

2. Write a dplyr pipe to change the units of age to "years" 
and circumference to "cm".

```{r q2}
Orange = Orange %>% 
  mutate(age = age / 365, 
         circumference = circumference / 10)
Orange %>% head()
```

 
3. Write a dplyr pipe to add a column assigning a
z-score to each tree, centered around the mean for all trees at a given age.

```{r q3}
Orange %>% 
  group_by(Tree) %>% 
  mutate(z_score = (age - mean(age)) / sd(age))
#? I can't really understand this question
```

 
4. Write a dplyr pipe to calculate the average rate of growth
(cm/year) across all trees between between age 0 (assume circumference = 0) and
the final measured age for each tree.

```{r q4}
Orange %>% 
  group_by(Tree) %>%
  filter(row_number() == n()) %>%
  mutate(growth_rate = circumference/age) %>%
  group_by() %>%
  summarise(avg_growth_rate = mean(growth_rate))
```

5. Describe the result of the following pipe in 1-3 sentences. Your
    *description* should touch on both the rows and columns and also describe
    a substantive question addressed by the result. 

```r
 Orange %>%
   group_by(Tree) %>%
   mutate( 
     new_growth = c(circumference[1], diff(circumference)),
     elapsed = c(age[1], diff(age))
   ) %>%
   group_by(age) %>% 
   summarize( 
     avg_rate = mean(new_growth / elapsed),
     se = sd( new_growth / elapsed ) / sqrt(n()) 
   )
```

**Description: **
For each one of the seven recorded age (118, 484, ..., 1582 days), 
compute the mean and standard error of growth rate
(the increase in circumference divided by the increase in age) 
across all trees.  
**Substantive Question: **
Is the fluctuation of average growth rates relevant to season? 
Because intuitively, the growth rate should be decreasing over time. 
But for these trees the `avg_rate` seems quite fluctuating. 

